---
title: "RCons Polling Station Data Update"
author: "Luke Sonnet"
date: "3/3/2019"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE)
```

This document will present a preliminary report of the RCons data entry of polling station level data for the 2018 General Elections in Pakistan. First let me load some packages and create a helper function.

```{r fns}
library(tidyverse)
library(haven)

find_non_coercable_chars <- function(x) {
  suppressWarnings(unique(x[!is.na(x) & is.na(as.numeric(x))]))
}
```

## Data submitted

The raw data are in 5 files by province:

* Form 49 (the data on the candidate names and votes, candidate_id to merge with other variables, constituency totals, etc)
* Form 48 (the polling station votes per candidate, total votes, valid votes, and more)
* Form 45 (unofficial polling station level results, only male and female vote totals were recorded)
* Form 28 - Province (list of census block-polling station pairs for the provincial assemblies)
* Form 28 - National (list of census block-polling station pairs for the provincial assemblies)

I'll load each in turn and discuss any irregularities.

**For RCons, please look at every line with the tag "RCONS". Please respond to each issue listed via email when you've had a chance to investiage and solve them. A report in email with any updated datasets would be perfect.**

## Form 49s

```{r f49}
f49_files <- list.files(
  "source_data/rcons_data",
  pattern = "49",
  recursive = TRUE,
  full.names = TRUE
)
f49_df <- map_dfr(f49_files, read_dta)

# RCONS: However, party names aren't standardized, would be helpful to have them standardized
# Here are some examples:
sort(unique(f49_df$party_affiliation))[40:60]

# What constituencies are there?
all_constituency_ids <- c(
  paste0("NA", 1:272),
  paste0("PS", 1:130),
  paste0("PP", 1:297),
  paste0("PK", 1:99),
  paste0("PB", 1:51)
)

# Quite a few constituencies missing from RCons data
rcons_missing <- setdiff(all_constituency_ids, f49_df$constituency_id)
rcons_missing
# RCONS: How many of these constituencies had elections postponed, how many are missing
# for other reasons?

glimpse(f49_df)
```

## Form 48s

```{r f48}
f48_files <- list.files(
  "source_data/rcons_data",
  pattern = "48",
  recursive = TRUE,
  full.names = TRUE
)

f48_temp_df <- map_dfr(
  f48_files,
  ~ {
    read_dta(.x) %>%
      mutate_at(
        vars(contains("_votes")),
        funs(as.character)
      ) %>%
      mutate(file = gsub("source_data/rcons_data/", "", .x))
  }
)

# RCONS: error in some of the votes data, + symbols, duplication of -99s, etc.
f48_temp_df %>%
  filter_at(
    vars(contains("_votes")),
    any_vars(grepl("\\+|/|\\-99\\-99$", .))
  ) %>%
  select(constituency_id,
         starts_with("ps_id"),
         contains("_votes"),
         -matches("can_votes_(19|[234][0-9]+)"),
         file) %>%
  as.data.frame()

# All constituency_id, ps_id combos are unique!
any(duplicated(select(f48_temp_df, constituency_id, ps_id)))
```

## Form 45s

```{r f45}
f45_files <- list.files(
  "source_data/rcons_data",
  pattern = "45",
  recursive = TRUE,
  full.names = TRUE
)

f45_temp_df <- map_dfr(
  f45_files,
  ~ {
    read_dta(.x) %>%
      mutate_at(
        vars(contains("_turnout"), contains("total_votes")),
        funs(as.character)
      ) %>%
      mutate(file = gsub("source_data/rcons_data/", "", .x))
  }
)

# RCONS: error in some of the turnout and votes data, + symbols
f45_temp_df %>%
  filter_at(
    vars(contains("_turnout"), contains("total_votes")),
    any_vars(grepl("\\+", .))
  ) %>%
  select(constituency_id, starts_with("ps_id"), contains("turnout"), total_votes) %>%
  as.data.frame()

# RCONS: also many duplicates of rows, is this expected?
any(duplicated(select(f45_temp_df, constituency_id, ps_id)))
f45_temp_df %>%
  group_by(constituency_id, ps_id) %>%
  filter(n() > 1) %>%
  select(constituency_id, ps_id, file) %>%
  arrange(constituency_id, ps_id)
```

## Form 28s

```{r f28}
f28_files <- list.files(
  "source_data/rcons_data",
  pattern = "Polling_Station_List|28",
  recursive = TRUE,
  full.names = TRUE
)

# Just read in to list
f28s <- map(f28_files, read_dta)

# Errors in `block_code_rural` and `block_code_urban`
# 1) RCONS: The the rural and urban block codes in Sindh have the dashes in them. Is this an error or is this how they appear on the forms?
map(f28s, ~{find_non_coercable_chars(.x$block_code)[1:10]})[7:8]
# 2) RCONS: Typo in  block_code_rural in the Punjab PA data
f28s[[3]] %>%
  filter(block_code_rural %in% c("26030907 (Part\n2)", "m   81021103")) %>%
  select(constituency_id, starts_with("ps_id"), starts_with("block_code"))

# RCONS: `male_booths`, `female_booths`, and `total_booths` have some errors
# in the KP provincial data
f28s[[4]] %>%
  filter(male_booths %in% c("4\n4", "Male")) %>%
  select(constituency_id, starts_with("ps_id"), ends_with("_booths"))

# RCONS: Plus signs in male_voters codes for Punjab PA and NA;
# there are plus signs in lots of the data later on
f28s[[5]] %>%
  filter(grepl("\\+", male_voters)) %>%
  select(constituency_id, starts_with("ps_id"), male_voters)

# If we hide the above problems we can get the data anyways
f28_df <- map_dfr(
  f28_files,
  ~ {
    read_dta(.x) %>% 
      mutate_at(
        vars(starts_with("block_code")), funs(gsub("null", NA, as.character(.)))
      ) %>%
      mutate(male_voters = gsub("\\+", "", male_voters)) %>%
      mutate_at(vars(ends_with("_booths")), funs(as.numeric)) 
  }
)

glimpse(f28_df)
```
